MSPMS Report

### 2024-07-25 README.Rmd
Incorporated new qc check functionality

There are 4 different types of functions in this package.

@@

2. Data processing/ normalization. These functions allow the user to normalize
and process the MSP-MS data.

3. Statistics. These methods allow the user to perform basic statistics on the 
normalized/processed data.

4. Data visualization. These functions allow the user to visualize the data.

@@

@@

>>>>

There are 5 different types of functions in this package.

@@

2. QC checks. These functions allow the user to conduct quality control checks

3. Data processing/ normalization. These functions allow the user to normalize
and process the MSP-MS data.

4. Statistics. These methods allow the user to perform basic statistics on the 
normalized/processed data.

5. Data visualization. These functions allow the user to visualize the data.

@@

**QC checks**
1. *qc_check()*: Conducts quality control checks on the data.
2. *plot_qc_check()*: Plots the results of the qc_checks() function.
3. *find_nd_peptides()*: Finds library peptides that were not detected.
4. *plot_nd_peptides()*: Plots the results of the find_nd_peptides() function.

@@

## QC Checks

In theory, all of the full length peptides present in the library 
should be detected in the MSP-MS experiment. In practice, this isn't the case
because some peptides fly worse than others.

The check_qc() function let's us know what percentage of these full length
peptides are missing. This function also checks to see what

We can use it as follows:

```{r}
qc_check <- mspms::qc_check(prepared_data = mspms::peaks_prepared_data,
                            peptide_library = mspms::peptide_library,
                            design_matrix = mspms::design_matrix)

head(qc_check)
```

These results can then be visualized as follows using plot_qc_check()

First lets look at the percent of full length peptides in the library detected

```{r}
plot_qc_check(qc_check)[1]
```

We can see here that most samples have more than 90% of the full length peptides
detected. We would expect to see ~90% of the peptides detected at T0. Notably 
less than this would suggest that something funky happened to the sample and it

all of these samples look reasonably good.

Now let's look at all the percentage of peptides in the library detected, 
including both full length and detected fragments.

```{r}
plot_qc_check(qc_check)[2]
```

Here we see that most samples have more than 90% of the peptides detected. 
This is as would be expected with quality data, thus Ww can consider these 
samples to have passed QC. Note that we see more peptides here compared to the 
previous qc plot, since some of the digested peptide fragments fly better than
the parent peptides on the mass spec.

It may be helpful to know what peptides from the library were not detected in a 
given experiment We can examine this as follows:

```{r}
nd_peptides = mspms::find_nd_peptides(mspms::peaks_prepared_data,
                                      mspms::peptide_library,
                                      mspms::design_matrix)

head(nd_peptides)
```

we can then visualize as follows using the plot_nd_peptides() function.

```{r}
plot_nd_peptides(nd_peptides)
```
Here we can see that some of the full length peptides are not detected in 100% 
of samples. Additionally, some of these library peptides are not detected, even
when considering possible fragment peptides.

### 2024-07-25 inst/rmarkdown/templates/mspms_report/skeleton/skeleton.Rmd
Incorporated new qc check functionality

@@

**Figure 1.** Here a PCA of the experiment is displayed. The color represents time, while shape and linetype of eclipse (if there are enough points to calculate) represents condition.

@@

**Figure 2.** Here an interactive heatmap of the experiment is displayed. Mouse over to see more information on each cell. Column side colors represent condition and time.

@@

**Table 1. ** Here anova results are displayed for each condition (testing to see if there is an effect on time).

```{r}
t2 = log2fc_t_test(mspms_data) %>% 
  dplyr::arrange(p.adj)

@@

**Table 2.** Here T-test results are displayed as compared to time 0 for each condition

```{r}
plot_volcano(t2)
```

**Figure 3.** Here the Log2fc and T tests are displayed as volcano plots relative to T0 for each condition.

@@

```{r,fig.height = 10, fig.width = 6}
print(mspms::plot_all_icelogos(mspms_data))
```

**Figure 4.** Here the cleavage motifs enriched in each condition are shown. 
Enriched cleavage motifs are defined as having a p.adj <= 0.05 and a log2fc >3
relative to time 0.

@@

d = count_cleavages_per_pos(sig)

plot_cleavages_per_pos(d)
```

**Figure 5.** Here the count of significant cleavages at each position of the 
peptide library are shown for each condition (facet) per time point (color). 
Significant cleavages are defined as having a p.adj <= 0.05 and a log2fc > 3 
relative to time 0.

>>>>

# QC Checks

```{r}
qc_checks = mspms::qc_check(prepared_data,peptide_library,design_matrix)

DT::datatable(qc_checks)

qc_checks %>%
  downloadthis::download_this(
    output_name = "qc_checks",
    output_extension = ".csv",
    button_label = "Download qc_check",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save",
    csv2 = FALSE
  )
```

**Table 1.** Here the percentage of the peptide library that is detected in 
each sample is displayed.

```{r}
mspms::plot_qc_check(qc_checks)

```

**Figure 1.** Here the results of the qc checks are displayed.

```{r}
nd_peptides = mspms::find_nd_peptides(prepared_data,
                                      peptide_library,
                                      design_matrix)

DT::datatable(nd_peptides)

nd_peptides %>%
  downloadthis::download_this(
    output_name = "nd_peptides",
    output_extension = ".csv",
    button_label = "Download nd_peptides",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save",
    csv2 = FALSE
  )
```

**Table 2.** Here the peptides missing from each sample are shown

```{r}
plot_nd_peptides(nd_peptides)
```

**Figure 2.** Here the library peptides not detected in each samples are shown.

@@

**Figure 3.** Here a PCA of the experiment is displayed. The color represents time, while shape and linetype of eclipse (if there are enough points to calculate) represents condition.

@@

**Figure 4.** Here an interactive heatmap of the experiment is displayed. Mouse over to see more information on each cell. Column side colors represent condition and time.

@@

**Table 3. ** Here anova results are displayed for each condition (testing to see if there is an effect on time).

```{r}
t2 = mspms::log2fc_t_test(mspms_data) %>% 
  dplyr::arrange(p.adj)

@@

**Table 4.** Here T-test results are displayed as compared to time 0 for each condition

```{r}
mspms::plot_volcano(t2)
```

**Figure 5.** Here the Log2fc and T tests are displayed as volcano plots relative to T0 for each condition.

@@

```{r,fig.height = 10, fig.width = 6}

tryCatch({
  print(mspms::plot_all_icelogos(mspms_data))
}, error = function(e) {
  print(e)
})
```

**Figure 6.** Here the cleavage motifs enriched in each condition are shown. 
Enriched cleavage motifs are defined as having a p.adj <= 0.05 and a log2fc >3
relative to time 0.

@@

d = mspms::count_cleavages_per_pos(sig)

tryCatch({
  print(mspms::plot_cleavages_per_pos(d))
}, error = function(e) {
  print(e)
})
```

**Figure 7.** Here the count of significant cleavages at each position of the 
peptide library are shown for each condition (facet) per time point (color). 
Significant cleavages are defined as having a p.adj <= 0.05 and a log2fc > 3 
relative to time 0.
